---
title: "Using gamar"
author: "Marc Choisy, Nicolas Marilleau, Jean-Daniel Zucker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=F}
knitr::knit_hooks$set(margin = function(before,options,envir) {
  if(before) par(mgp=c(1.5,0.5,0),bty="n",plt=c(.105,.97,.13,.97))
  else NULL
})

knitr::opts_chunk$set(margin=T,prompt=T,comment="",collapse=T,cache=T,
                      dev.args=list(pointsize=11),fig.height=3.5,
                      fig.width=4.24725,fig.retina=2,fig.align="center")

dir.create("tmp")
knitr::opts_knit$set(root.dir=paste0(getwd(),"/tmp"))
```

This vignette shows how to install and use the `gamar` package.

## Installing `gamar`

The package is currently on the GitHub repository. To install it from this
repository you can use the `install_github` function from the `devtools`
package. If this package is not installed on your system, install it

```{r eval=F}
install.packages("devtools")
```

Then you can install the `gamar` package by typing:

```{r eval=F}
devtools::install_github("choisy/gamar")
```

Note that the `gamar` package depends on the `XML` package. The installation of
`gamar` package automatically installs the `XML` too, if not already installed
on your system. Then, you can load and attach the `gamar` library:

```{r}
library(gamar)
```

Note that you need the GAMA platform (version > 1.7.0) to be installed on your
machine. For now, you have to set the path to the GAMA folder yourself:

```{r eval=F}
defpath("path_to_your_GAMA_folder")
```

On most Mac it would be:

```{r}
defpath("/Applications/Gama.app")
```

There are a number of example files in the `examples` directory of the library:

```{r}
(f <- system.file("examples",package="gamar"))
dir(f)
```

Move the GAML and XML files of the SIR model and experiment to the working
directory:

```{r}
dir()
file.copy(paste0(f,"/sir.gaml"),".")
dir()
```

Loading the `sir` experiment defined in the `sir.gaml` file:

```{r}
experimentclone1 <- getmodelparameter("sir.gaml","sir")
```

The variable monitored during the simulation:

```{r}
getoutputnames(experimentclone1)
```

The parameters of the model:

```{r}
getparameternames(experimentclone1)
```

Setting the values of the parameters (i.e. initial values of state variables
and model paramters):

```{r}
experimentclone1 <- setparametervalue(experimentclone1,"S0",990)
experimentclone1 <- setparametervalue(experimentclone1,"I0",10)
experimentclone1 <- setparametervalue(experimentclone1,"R0",0)
experimentclone1 <- setparametervalue(experimentclone1,"beta",.3)
experimentclone1 <- setparametervalue(experimentclone1,"gamma",.1)
```

Setting the frame rate of the monitored variable:

```{r}
experimentclone1 <- setoutputframerate(experimentclone1,"susceptibles",1)
```

Setting the number of steps to simulate:

```{r}
experimentclone1 <- setfinalstep(experimentclone1,2000)
```

Making a second experiment, modified from the first experiment (with new
starting values of susceptibles and infected):

```{r}
experimentclone2 <- setparametervalue(experimentclone1,"S0",995)
experimentclone2 <- setparametervalue(experimentclone2,"I0",5)
```

Putting the two experiment in one single experiment plan:

```{r}
experimentplan <- addtoexperimentplan(experimentclone1)
experimentplan <- addtoexperimentplan(experimentplan,experimentclone2)
```

Running the experiment plan on one core:

```{r}
outfiles <- startexperimentplan(experimentplan,hpc=2)
```




```{r include=F, cache=F}
setwd("..")
knitr::opts_knit$set(root.dir=".")
```

```{r include=F, cache=F}
unlink("tmp",T,T)
```
