---
title: "Using gamar"
author: "Marc Choisy, Nicolas Marilleau, Jean-Daniel Zucker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=F}
knitr::knit_hooks$set(margin = function(before,options,envir) {
  if(before) par(mgp=c(1.5,0.5,0),bty="n",plt=c(.105,.97,.13,.97))
  else NULL
})

knitr::opts_chunk$set(margin=T,prompt=T,comment="",collapse=T,cache=T,
                      dev.args=list(pointsize=11),fig.height=3.5,
                      fig.width=4.24725,fig.retina=2,fig.align="center")

dir.create("tmp")
knitr::opts_knit$set(root.dir=paste0(getwd(),"/tmp"))
```

This vignette shows how to install and use the `gamar` package.

## Installing `gamar`

The package is currently on the GitHub repository. To install it from this
repository you can use the `install_github` function from the `devtools`
package. If this package is not installed on your system, install it

```{r eval=F}
install.packages("devtools")
```

Then you can install the `gamar` package by typing:

```{r eval=F}
devtools::install_github("choisy/gamar")
```

Note that the `gamar` package depends on the `XML` package. The installation of
`gamar` package automatically installs the `XML` too, if not already installed
on your system. Then, you can load and attach the `gamar` library:

## Using `gamar`

```{r}
library(gamar)
```

Note that you need the GAMA platform (version > 1.7.0) to be installed on your
machine. For now, you have to set the path to the GAMA folder yourself:

```{r eval=F}
defpath("path_to_your_GAMA_folder")
```

On most Mac it would be:

```{r}
defpath("/Applications/Gama.app")
```

There are a number of example files in the `examples` directory of the library:

```{r}
(f <- system.file("examples",package="gamar"))
dir(f)
```

## An SIR model

Loading the `sir` experiment defined in the `sir.gaml` file:

```{r}
dir()
experiment1 <- getmodelparameter(paste0(f,"/sir.gaml"),"sir")
```

Note that the function `getmodelparameter` loads GAMA and creates the
`workgamar` folder:

```{r}
dir()
```

This folder contains at this stage an XML file containing the specifications of
the newly created experiment:

```{r}
dir("workgamar")
```

The parameters of the model:

```{r}
getparameternames(experiment1)
```

The variables monitored during the simulation:

```{r}
getoutputnames(experiment1)
```

Setting the values of the parameters (i.e. initial values of state variables
and model paramters):

```{r}
experiment1 <- setparametervalue(experiment1,"S0",990)
experiment1 <- setparametervalue(experiment1,"I0",10)
experiment1 <- setparametervalue(experiment1,"R0",0)
experiment1 <- setparametervalue(experiment1,"beta",.3)
experiment1 <- setparametervalue(experiment1,"gamma",.1)
```

Setting the frame rates of the monitored variables:

```{r}
experiment1 <- setoutputframerate(experiment1,"susceptibles",1)
experiment1 <- setoutputframerate(experiment1,"infected",1)
experiment1 <- setoutputframerate(experiment1,"recovered",1)
```

Setting the number of steps to simulate:

```{r}
experiment1 <- setfinalstep(experiment1,200)
```

Making a second experiment, modified from the first experiment (with new
starting values of susceptibles and infected):

```{r}
experiment2 <- setparametervalue(experiment1,"S0",995)
experiment2 <- setparametervalue(experiment2,"I0",5)
```

Putting the two experiment in one single experiment plan:

```{r}
experimentplan <- addtoexperimentplan(experiment1)
experimentplan <- addtoexperimentplan(experimentplan,experiment2)
```

Running the experiment plan on two cores:

```{r}
dir("workgamar")
output <- runexpplan(experimentplan,2)
```

Note the creation of another XML file and a folder

```{r}
dir("workgamar")
```

This later folder contains one XML file per experiment, containing the
experiment's output, and another folder:

```{r}
dir("workgamar/out_sir_1")
```

This later folder may contain the snapshots of the simulations.

At the end, we can also remove this `workgamar` folder:

```{r}
dir()
cleaning()
dir()
```

Plotting the dynamics of infected of the first experiment of the plan:

```{r}
with(output[[1]],{
  plot(step,I,type="l",lwd=2,col="red",
       ylab="number of infected",ylim=c(0,1000))
  lines(step,S,lwd=2,col="blue")
  lines(step,R,lwd=2,col="green")
})
```

## A Lotka-Volterra model

Loading the experiment `prey_predator` from the `predator_prey` model:

```{r}
experiment1 <- getmodelparameter(
  paste0(system.file("examples",package="gamar"),
         "/predator_prey/models/predator_prey.gaml"),"prey_predator")
```

The parameters of the model:

```{r}
getparameternames(experiment1)
```

The monitored variables of the model:

```{r}
getoutputnames(experiment1)
```

Setting the model parameters to values:

```{r}
experiment1 <- setparametervalue(experiment1,"Initial number of preys: ",990)
experiment1 <- setparametervalue(experiment1,"Initial number of predators: ",100)
experiment1 <- setparametervalue(experiment1,"Predator probability reproduce: ",0.1)
```

Specifying the number of steps:

```{r}
experiment1 <- setfinalstep(experiment1,100)
```

Adding the experiment to an experiment plan:

```{r}
experimentplan <- addtoexperimentplan(experiment1)
```

Running the experiment (only one here) of the plan:

```{r}
output <- runexpplan(experimentplan,1)
```

Plotting the dynamics of the number of preys:

```{r}
with(output[[1]],plot(step,`Number of preys`,type="l",lwd=2,col="red",
                      ylim=c(0,1050)))
```

We can also generate a movie of the simulation of one experiment:

```{r}
dir()
makemovie(output[[1]])
dir()
```


```{r include=F, cache=F}
setwd("..")
knitr::opts_knit$set(root.dir=".")
```

```{r include=F, cache=F}
unlink("tmp",T,T)
```
